### 2024-06-18 docs/ManuscriptGrp10.qmd

---
title: "Honey Bee Colonies Impacted by Varroa, American Fouldbrood and Global Warming"
subtitle: ""
author: Edward Cruz, Jr, Leonel G. Salazar, Amelia Tristan, Mojisola Popoola
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/apa.csl
---

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Summary/Abstract

Group Ten is pursuing the historical data collected by various public agencies to determine if the impact of Varroa, American Foulbrood, and global warming has demonstrated an increase in hive losses across the United States and Texas. The research will be accomplished using data sets derived from the agencies National Agricultural Statistics Service, Agricultural Statistics Board, and United States Department of Agriculture (USDA). Data sets containing several years of hive losses based on varroa and bacterium losses including recent possible environmental thermal global warming. A visualization of outcomes using R demonstrating and validating possible detrimental effects on honeybee colonies in the United States and Texas wrought by the negative impact of mites, bacterium, and global warming that will affect honey production and inevitably impact food production.

{{< pagebreak >}}

# Introduction

Honeybees have been nature's pollinators for centuries, with documented relationships with humans dating back to ancient Egyptians and Hindus. Historically, humans have maintained beehives, using honey as medicine in cultures such as the Egyptians, Assyrians, Chinese, Greeks, and Romans. The natural antibacterial properties of honey made it a valuable treatment for wounds, preventing infection, a practice used by Romans and Russians during World War I. Honeybees and other pollinators are critical for food production and nutritional security, yet bees face a variety of survival challenges. Currently, Varroa mites impact bee colonies and this mite infestation, a tiny red-brown parasite that can live on adult honeybees and reproduce on larvae and pupae in the developing brood. Another major threat is American Foulbrood Disease (AFB), caused by the bacterium Paenibacillus larvae. A disease that is fatal to honeybee larvae and found worldwide. The only effective control measure is to incinerate and destroy infected hives and live bees mitigating the infectious spread to other colonies. In addition, the exploration of climate change impacting honeybee colony losses has only recently been researched. While there are correlations between higher winter temperatures and greater colony losses, the effects of warmer autumn and winter temperatures on colony population dynamics and age structure as potential causes of reduced colony survival have not yet been fully investigated.Index Catalog // USDA Economics, Statistics and Market Information System. (n.d.-b). Index Catalog // USDA Economics, Statistics and Market Information System. (n.d.-a). , USDA - National Agricultural Statistics Service - Surveys - honey bee surveys and reports. (n.d.).\
<https://usda.library.cornell.edu/catalog?f%5Bkeywords_sim%5D%5B%5D=honey+bees&locale=en> <https://www.nass.usda.gov/Surveys/Guide_to_NASS_Surveys/Bee_and_Honey/>

## General Background Information

## Description of data and data source

Bee colonies maintained by beekeepers are considered livestock by the USDA due to their ability to produce honey, a consumable food item, and their essential role in assisting farmers with pollination crop seasons. Given the importance of bee colonies in agriculture, it was logical to source data from the following two authoritative websites: 1. USDA National Agricultural Statistics Service (NASS): This site provides comprehensive agricultural data, including statistics on honey production and colony health. 2. Bee Informed Partnership: This site offers detailed insights and research on bee colony management and health, contributing valuable information on the status and trends of bee populations. Index Catalog // USDA Economics, Statistics and Market Information System. (n.d.-a). https://usda.library.cornell.edu/catalog?f%5Bkeywords_sim%5D%5B%5D=honey+bees&locale=en\
USDA - National Agricultural Statistics Service - Surveys - honey bee surveys and reports. (n.d.). https://www.nass.usda.gov/Surveys/Guide_to_NASS_Surveys/Bee_and_Honey/

## Questions/Hypotheses to be addressed

Hypothesis: "The negative impacts of mites, bacterium, and global warming have detrimental effects on honeybee colonies in the United States and Texas, which in turn will lead to a decline in honey production and negatively impact food production." This hypothesis can be tested and validated through a visualization of outcomes using R, demonstrating the relationship between these factors and their effects on honeybee colonies.

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].

{{< pagebreak >}}

# Methods

*Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement.*

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender. We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
knitr::include_graphics(here("assets","antigen-recognition.png"))
```

## Data aquisition

We got our data from the United States Department of Agriculture (USDA).

## Data import and cleaning

We decided to clean out our data from a few different datasets. We had to remove blank spaces and columns that were not pertinent to our analysis. We then filtered out other observations that did not directly deal with the data we are exploring. We are looking for cause of death to bee colonies and how they are affected by mites and climate change so we wanted to single out data that represented the losses so we can explore the different states by year and determine how the colonies were affected.

```{r}
library(readxl)
library(tidyverse)

```

```{r}

data <- read.csv("C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/docs/DatabyState.csv")

View(data)
```

```{r}

# Select all columns except 3, 6, and 9
Data_Clean <- dplyr::select(data, -c(3, 6, 9))

```

```{r}

# Output cleaned data file to a csv file.
write.csv(Data_Clean, "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/docs/Databystate_Clean.csv")

view(Data_Clean)

```

```{r}
library(readxl)
library(tidyverse)

```

```{r}

data <- read.csv("C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/docs/DatabyState.csv")

```

```{r}

# Select all columns except 3, 6, and 9
Data_Clean <- dplyr::select(data, -c(3, 6, 9))

```

```{r}

# Output cleaned data file to a csv file.
write.csv(Data_Clean, "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/docs/Databystate_Clean.csv")

```

```{r}

# Assuming your data frame is named "data"

# Filter for rows where "Loss" or "Deadout" is present in any column (case-insensitive)
library(stringr)  # Load stringr package for regular expressions
data_filtered <- Data_Clean[rowSums(sapply(data, grepl, pattern = c("Loss"), ignore.case = TRUE)) > 0, ]

```

```{r}

# Output cleaned data file to a csv file.
write.csv(Data_Clean, "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/docs/Databystate_Filtered.csv")

```

```{r}
ggplot(data_filtered, aes( Data.Item, State.ANSI)) + geom_boxplot()
```

```{r}
ggplot(data_filtered, aes(Year)) + geom_histogram()
```

```{r}
ggplot(data_filtered, aes(x = Year)) + 
  geom_histogram(bins = 10, fill = "blue", color = "black", alpha = 0.7) +  # Adding fill color, border color, and transparency
  labs(title = "Histogram of Year", x = "Year", y = "Count") +  # Adding labels
  theme_minimal() 
```

```{r}
library(readxl)
library(tidyverse)
library(dplyr)

```

```{r}
# Output cleaned data file to a csv file.
hcny_data <- read.csv("C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/docs/hcny_CleanDraft.csv", header = FALSE, stringsAsFactors = FALSE)

#view(hcny_data)
```

```{r}
# Read the CSV file
hcny_data <- read.csv("C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/docs/hcny_CleanDraft.csv", header = FALSE, stringsAsFactors = FALSE)

```

```{r}
# Merge the first two rows to create a proper header
header <- hcny_data[1:2, ]
header <- sapply(header, function(x) paste(na.omit(x), collapse = " "))
colnames(hcny_data) <- header

# Remove the first two rows (header rows) and any completely blank rows
data_clean2 <- hcny_data[-c(1, 2), ]

# Remove rows that are completely blank
data_clean2 <- data_clean2[rowSums(is.na(data_clean2) | data_clean2 == "") != ncol(data_clean2), ]

# Convert columns to numeric where applicable (if they contain percentage values)
hcnydata_cleaned <- data_clean2 %>%
  mutate(across(.cols = -1, .fns = ~ as.numeric(gsub("[^0-9.-]", "", .))))

# Assuming hcny_data_cleaned is your cleaned data frame
colnames(hcnydata_cleaned) <- c("state", "varroa_mites", "other_pests", "disease", "pesticides", "other", "unknown")
```

```{r}
# Write the cleaned data to a new CSV file
write.csv(hcnydata_cleaned, "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/docs/hcnydata_cleaned.csv", row.names = FALSE)

# View the updated data frame
view(hcnydata_cleaned)
```

```{r}
ggplot(hcnydata_cleaned, aes(varroa_mites)) + geom_histogram()
```

## Statistical analysis

*Explain anything related to your statistical analyses.*

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

*Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project.*

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below. I generally recommend the `here` package.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/tables/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

*To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
```

## Full analysis

*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here.*

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

*Summarize what you did, what you found and what it means.*

## Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your analysis.*

## Conclusions

*What are the main take-home messages?*

*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like.

{{< pagebreak >}}

# References