### 2024-06-21 docs/coding-exercise/coding-exerciseECruzkml614.qmd
Leo,
3rd time is a charm, I believe that my file coding-exerciseECruzkml614.qmd can be found in my docs/coding-exercise folder

---
title: "R Coding Exercise"
author: "Edward Cruz kml614"
date: "June 18, 2024"
execute: 
  warning: false
  error: true
format:
  html:
    theme: ""
---

## Loading and Checking Data

### Import libraries

```{r}
update.packages('dslabs')
library(dslabs) # to import dataset
library(tidyverse) # data aggregation
library(dplyr) # data aggregation? 
data(gapminder)
```

### Help Function

```{r}
help("gapminder")
```

### Data Structure Overview

```{r}
str(gapminder)
```

### Data Summary

```{r}
summary(gapminder)
```

### Class Functions - Object Determination

```{r}
class(gapminder)
```

## Processing Data

### Assign only African countries to a new object/variable.

```{r}
africadata = filter(gapminder, continent == 'Africa')
head(africadata)
```

### Africa Data Structure Overview

```{r}
str(africadata)
```

### Africa Summary

```{r}
summary(africadata)
```

### African Infant Mortality Object

```{r}
infant_survival_rate = africadata %>%
  select(infant_mortality, life_expectancy)
head(infant_survival_rate)
```

### African Infant Mortality Data Structure

```{r}
str(infant_survival_rate)
```

### African Infant Mortality Data Summary

```{r}
summary(infant_survival_rate)
```

### African Mortality Object

```{r}
african_mortality = africadata %>%
  select(population, life_expectancy)
head(african_mortality)
```

### African Mortality Data Structure

```{r}
str(african_mortality)
```

### African Mortality Data Summary

```{r}
summary(african_mortality)
```

## Plotting

### Life Expectancy as a Function of Infant Mortality

```{r}
regression = lm(infant_survival_rate$life_expectancy ~ infant_survival_rate$infant_mortality)
rsq <- function(x, y) summary(lm(infant_survival_rate$life_expectancy ~ infant_survival_rate$infant_mortality))$r.squared
r_square = round(rsq(obs, mod), 3)
# summary(regression)$coefficients - keep to sanity check
coeff = round(summary(regression)$coefficients[2],3)
y_int = round(summary(regression)$coefficients[1],3)
plot(
  infant_survival_rate$infant_mortality,
  infant_survival_rate$life_expectancy,
  main='African Life Expectancy \n Infant Mortality Effect',
  xlab='Infant Mortality',
  ylab='Life Expectancy',
   col='aquamarine', 
   bg='cornflowerblue', 
   pch = 21,
)
# rsq_str = expression(paste("r"^"2:"))
rsq_str = "r2: "
text(175, 75, paste0('y = ', coeff, 'x + ', y_int), cex=0.75)
text(170, 70, paste0(rsq_str, r_square), cex=0.75)
abline(
  lm(
    infant_survival_rate$life_expectancy ~ infant_survival_rate$infant_mortality),
  col='darkblue',
  lwd=1.75)
```

### Life Expectancy as a Function of Population

```{r}
regression = lm(african_mortality$life_expectancy ~ african_mortality$population)
rsq <- function(x, y) summary(lm(african_mortality$life_expectancy ~ african_mortality$population))$r.squared
r_square = round(rsq(obs, mod), 3)
# summary(regression)$coefficients - keep to sanity check
coeff = round(summary(regression)$coefficients[2],9)
y_int = round(summary(regression)$coefficients[1],3)
plot(
  african_mortality$population,
  african_mortality$life_expectancy,
  main='African Life Expectancy \n Population Effect',
  xlab='Population',
  ylab='Life Expectancy',
  col='aquamarine', 
  bg='cornflowerblue', 
  pch = 21,
)
# rsq_str = expression(paste("r"^"2:"))
rsq_str = "r2: "
text(155000000, 75, paste0('y = ', coeff, 'x + ', y_int), cex=0.75)
text(155000000, 70, paste0(rsq_str, r_square), cex=0.75)
abline(
lm(
  african_mortality$life_expectancy ~ african_mortality$population),
col='darkblue',
lwd=1.75)
```

\*\* STILL NEED TO ANSWER QUESTION AT THE END OF THE PLOTTING SECTION\*\*

## More Data Processing

### Missing Years For Infant Mortality Data

```{r}
miss_inf_mort_rows = africadata %>%
  filter(is.na(infant_mortality))
head(miss_inf_mort_rows)
```

### Perform Year Groupby to Identify Distinct Missing Years

```{r}
miss_inf_mort_yrs = miss_inf_mort_rows %>%
  group_by(year) %>%
  count(infant_mortality)
miss_inf_mort_yrs
```

### Extract Only Year 2000 From africadata As New Object

```{r}
africadata2000 = filter(africadata, year == 2000)
africadata2000
```

### africadata2000 Data Structure

```{r}
str(africadata2000)
```

### africadata2000 Data Summary

```{r}
summary(africadata2000)
```

## More Plotting

### Create The Same Plots For africadata2000 Object

#### Filter New Object to Create Graph Datasets (1/2)

```{r}
african_mortality2000 = africadata2000 %>%
  select(population, life_expectancy)
head(african_mortality2000)
```

#### Filter New Object to Create Graph Datasets (2/2)

```{r}
infant_survival_rate2000 = africadata2000 %>%
  select(infant_mortality, life_expectancy)
head(african_mortality2000)
```

#### Infant Mortality Effect

```{r}
regression = lm(infant_survival_rate2000$life_expectancy ~ infant_survival_rate2000$infant_mortality)
rsq <- function(x, y) summary(lm(infant_survival_rate2000$life_expectancy ~ infant_survival_rate2000$infant_mortality))$r.squared
r_square = round(rsq(obs, mod), 3)
# summary(regression)$coefficients - keep to sanity check
coeff = round(summary(regression)$coefficients[2],3)
y_int = round(summary(regression)$coefficients[1],3)
plot(
  infant_survival_rate2000$infant_mortality,
  infant_survival_rate2000$life_expectancy,
  main='African Life Expectancy \n Infant Mortality Effect',
  xlab='Infant Mortality',
  ylab='Life Expectancy',
   col='aquamarine', 
   bg='cornflowerblue', 
   pch = 21,
)
# rsq_str = expression(paste("r"^"2:"))
rsq_str = "r2: "
text(130, 75, paste0('y = ', coeff, 'x + ', y_int), cex=0.75)
text(130, 73, paste0(rsq_str, r_square), cex=0.75)
abline(
  lm(
    infant_survival_rate2000$life_expectancy ~ infant_survival_rate2000$infant_mortality),
  col='darkblue',
  lwd=1.75)
```

#### Population Effect

```{r}
regression = lm(african_mortality2000$life_expectancy ~ african_mortality2000$population)
rsq <- function(x, y) summary(lm(african_mortality2000$life_expectancy ~ african_mortality2000$population))$r.squared
r_square = round(rsq(obs, mod), 3)
# summary(regression)$coefficients - keep to sanity check
coeff = round(summary(regression)$coefficients[2],9)
y_int = round(summary(regression)$coefficients[1],3)
plot(
  african_mortality2000$population,
  african_mortality2000$life_expectancy,
  main='African Life Expectancy \n Population Effect',
  xlab='Population',
  ylab='Life Expectancy',
  col='aquamarine', 
  bg='cornflowerblue', 
  pch = 21,
)
# rsq_str = expression(paste("r"^"2:"))
rsq_str = "r2: "
text(110000000, 75, paste0('y = ', coeff, 'x + ', y_int), cex=0.75)
text(110000000, 73, paste0(rsq_str, r_square), cex=0.75)
abline(
lm(
  african_mortality2000$life_expectancy ~ african_mortality2000$population),
col='darkblue',
lwd=1.75)
```

## Simple Model Fits

### Fit Life Expectancy to Infant Mortality

```{r}
fit1 = lm(life_expectancy ~ log(infant_mortality), data=infant_survival_rate2000)
summary(fit1)
```

```{r}
# Linear regression
fit1 = lm(life_expectancy ~ infant_mortality, data=infant_survival_rate2000)
summary(fit1)

# Polynomial regression of degree 2
fit2 = lm(life_expectancy ~ poly(infant_mortality, 2), data=infant_survival_rate2000)
summary(fit2)

# Plotting
plot(
  infant_survival_rate2000$infant_mortality,
  infant_survival_rate2000$life_expectancy,
  main='Infant Mortality vs Life Expectancy',
  xlab='Infant Mortality',
  ylab='Life Expectancy',
  col='aquamarine',
  pch=19
)

# Add linear regression line
abline(fit1, col='darkblue')

# Add polynomial regression line
new_data <- data.frame(infant_mortality = seq(min(infant_survival_rate2000$infant_mortality), max(infant_survival_rate2000$infant_mortality), length.out=100))
lines(new_data$infant_mortality, predict(fit2, new_data), col='cornflowerblue')

```

### Fit Life Expectancy to Population

```{r}
# Logarithmic regression
fit2 = lm(life_expectancy ~ log(population), data=african_mortality2000)
summary(fit2)

```

```{r}
# Linear regression
fit1 = lm(life_expectancy ~ population, data=african_mortality2000)
summary(fit1)

# Polynomial regression of degree 2
fit2 = lm(life_expectancy ~ poly(population, 2), data=african_mortality2000)
summary(fit2)

# Plotting
plot(
  african_mortality2000$population,
  african_mortality2000$life_expectancy,
  main='Population vs Life Expectancy',
  xlab='Population',
  ylab='Life Expectancy',
  col='aquamarine',
  pch=19
)

# Add linear regression line
abline(fit1, col='darkblue')

# Add polynomial regression line
new_data <- data.frame(population = seq(min(african_mortality2000$population), 
                                        max(african_mortality2000$population), length.out=100))
lines(new_data$population, predict(fit2, new_data), col='cornflowerblue')

# Add a legend to differentiate the lines
legend("topright", legend=c("Linear Fit", "Polynomial Fit (degree 2)"), col=c("darkblue", "cornflowerblue"), lwd=2)
```

#### Conclusion