### 2024-06-27 hcny_cleaned_updatedMoji.qmd

---
title: "loading/cleaning/wrangling/exploring"
author: "Mojisola Popoola, Leonel Salazar, ED Cruz, Amelia"
format: docx
---

```{r, message=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
library(dplyr)
library(readr)
library(tidymodels)
```

```{r}
# Output cleaned data file to a csv file.
hcny_data <- read.csv("C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/hcny_CleanDraft.csv", header = FALSE, stringsAsFactors = FALSE)
```

```{r}
head(hcny_data)
```

```{r}
str(hcny_data)
```

```{r}
summary(hcny_data)
```

```{r}
# Step 2: Merge the first two rows to create a new header
new_header <- paste(hcny_data[1, ], hcny_data[2, ], sep = " ")

# Step 3: Set the new header
colnames(hcny_data) <- new_header

# Step 4: Remove the first three rows
hcny_data <- hcny_data[-c(1:3), ]

# Step 5: Remove completely blank rows
hcny_data <- hcny_data[rowSums(hcny_data != "") > 0, ]

# Display the resulting data frame
head(hcny_data)
```

```{r}
# Assuming hcny_data is your cleaned data frame
colnames(hcny_data) <- c("state", "varroa_mites", "other_pests", "disease", "pesticides", "other", "unknown")
```

```{r}
head(hcny_data)
```

```{r}
# Convert specified columns to numeric with warning suppression
hcnydata_cleaned <- hcny_data %>%
  mutate(across(c("varroa_mites", "other_pests", "disease", "pesticides", "other", "unknown"), ~suppressWarnings(as.numeric(.))))

# Display the resulting data frame
head(hcnydata_cleaned)
```

```{r}
head(hcnydata_cleaned)
```

```{r}
# Verify the changes
str(hcnydata_cleaned)
```

```{r}
head(hcnydata_cleaned)
```

```{r,message=FALSE, warning=FALSE}
ggplot(hcnydata_cleaned, aes(varroa_mites)) + geom_histogram()
```

```{r}
view((hcnydata_cleaned))
```

```{r}
# Write the cleaned data to a new CSV file
write.csv(hcnydata_cleaned, "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/hcny_CleanDraft.csv", row.names = FALSE)
```

********************Conclusion- all code below is test code right now********************************

```{r}
# Assuming hcnydata_cleaned is your data frame
hcnydata_cleaned <- hcnydata_cleaned[apply(hcnydata_cleaned, 1, function(x) !all(is.na(x))), ]

# Print the modified data frame to verify the changes
head(hcnydata_cleaned)
view(hcnydata_cleaned)

```

```{r}
str(hcnydata_cleaned)
```

```{r, message=FALSE}
#update.packages('mgcv')
library(mgcv)
```

```{r}
# Convert the 'state' column to a factor
hcnydata_cleaned$state <- as.factor(hcnydata_cleaned$state)

# Display the resulting data frame
head(hcnydata_cleaned)
```

```{r}
# Step 3: Reindex the rows starting from 1
rownames(hcnydata_cleaned) <- NULL
```

```{r}
# Identify the rows to exclude
rows_to_exclude <- c(46)

# Remove rows 45 to 52 while excluding row 46
data_hcny <- hcnydata_cleaned[-c(45:52)[-which(c(45:52) %in% rows_to_exclude)], ]
```

```{r}
library(ggplot2)

# Create a bar graph to show state and varroa_mites
ggplot(data = data_hcny, aes(x = state, y = varroa_mites)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Varroa Mites by State",
       x = "State",
       y = "Varroa Mites Count") +
  theme_minimal() +
  coord_flip()
```

# Remove all rows with any NA values

```{r}

library(tidyr)

my_data <- data_hcny %>% drop_na()

# Verify that NAs have been removed
glimpse(my_data)
```
# Bivariate Analysis

```{r, message=FALSE, warning=FALSE}
## Example of a bivariate plot between 'state' and 'varroa_mites'
ggplot(my_data, aes(x = disease, y = varroa_mites)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Bivariate Relationship between State and Varroa Mites",
       x = "State",
       y = "Varroa Mites")
```
# Multivariable GLM

```{r}
install.packages("tidymodels")
library(tidymodels)
library(broom)

# Specify the GLM model
glm_spec <- linear_reg() %>%
  set_engine("glm")

# Create a recipe for preprocessing the data
glm_recipe <- recipe(varroa_mites ~ other_pests + disease + pesticides + other + unknown, data = my_data) %>%
  step_normalize(all_predictors())

# Create a workflow
glm_workflow <- workflow() %>%
  add_model(glm_spec) %>%
  add_recipe(glm_recipe)

# Fit the model
glm_fit <- fit(glm_workflow, data = my_data)

# Print the model summary using tidy()
model_summary <- tidy(glm_fit)
print(model_summary)
```