### 2024-07-04 cdcdata-exercise/deathcount_cleaned.qmd
Updated R Code to create synthetic data 07.04.24

---
title: "loading/cleaning/wrangling/exploring"
author: "Edward Cruz"
format: docx
---

```{r}
library(readxl)
#update.packages("tidyverse")
library(tidyverse)
#update.packages("dplyr")
library(dplyr)
#update.packages('mgcv')
library(mgcv)
```

```{r}
# Output cleaned data file to a csv file.
DthCount_data <- read.csv("C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/cdcdata-exercise/Weekly_DeathCounts.csv", header = FALSE, stringsAsFactors = FALSE)
```

```{r}
head(DthCount_data)
```

```{r}
str(DthCount_data)
```

```{r}
summary(DthCount_data)
```

```{r}
#view(DthCount_data)
```

```{r}
# Merge the first two rows to create a proper header
header <- DthCount_data[1:2, ]
header <- sapply(header, function(x) paste(na.omit(x), collapse = " "))
colnames(DthCount_data) <- header
head(header)
```

```{r}
# Remove the first row (header rows) and any completely blank rows
DthCount_cleaned <- DthCount_data[-c(1,1), ]
head(DthCount_cleaned)
```

```{r}
#view(DthCount_cleaned)
```

```{r}
# Assuming DthCount_cleaned is your cleaned data frame
colnames(DthCount_cleaned) <- c("Jurisdiction", "Week_Ending_Date", "State", "Year", "Week","Cause","Death_Count","Cause_Subgroup","Period","Suppress","Note","Avg_No_Deaths","Variance_2015_2019_2020", "Perc_Diff_2015_2019_2020","Types")
```

```{r}
head(DthCount_cleaned)
```

```{r}
# Convert specified columns to numeric
DthCount2_cleaned <- DthCount_cleaned %>%
  mutate(across(c("Year", "Week", "Death_Count", "Avg_No_Deaths", "Variance_2015_2019_2020", "Perc_Diff_2015_2019_2020"), as.numeric))
```

```{r}
str(DthCount2_cleaned)
```

```{r}
head(DthCount2_cleaned)
```

```{r}
# Write the cleaned data to a new CSV file
write.csv(DthCount2_cleaned, "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/cdcdata-exercise/DthCount2_cleaned.csv", row.names = FALSE)
```

```{r}
DthCount3_cleaned <- DthCount2_cleaned %>% select(-Suppress,-Note,-Types)

```

```{r}
#view(DthCount3_cleaned)
```

```{r}
ggplot(DthCount3_cleaned, aes(Death_Count)) + geom_histogram()
```

```{r}
ggplot(DthCount3_cleaned, aes(x = Avg_No_Deaths)) +
  geom_histogram(binwidth = 950, fill = "blue", color = "black") +
  labs(title = "Histogram of Death Count",
       x = "Average Number of Deaths",
       y = "Frequency")
```

```{r}
# Example scatter plot of two variables
ggplot(DthCount3_cleaned, aes(x = Death_Count, y = Avg_No_Deaths)) +
  geom_point() +
  labs(title = "Scatter Plot",
       x = "Deaths",
       y = "Average Number of Deaths")
```

```{r}
# Enhanced scatter plot of two variables
ggplot(DthCount3_cleaned, aes(x = Death_Count, y = Avg_No_Deaths)) +
  geom_point(color = "blue", size = 5, alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red", linetype = "dashed") +
  labs(title = "Scatter Plot of Death Count vs Average Number of Deaths",
       x = "Death Count",
       y = "Average Number of Deaths") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )
```

```{r}
# Bar chart of Death Count per Year
ggplot(DthCount3_cleaned, aes(x = Year, y = Death_Count)) +
  geom_bar(stat = "identity", outline = "black", color = "darkorange", alpha = .8) +
  labs(title = "Bar Chart of Death Count Per Year",
       x = "Year",
       y = "Death Count") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

```

```{r}
# Write the cleaned data to a new CSV file
write.csv(DthCount3_cleaned, "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/cdcdata-exercise/DthCount3_cleaned.csv", row.names = FALSE)
```

```{r}
# Assuming dat_cleaned is your data frame
DthCount3A_cleaned <- DthCount3_cleaned[apply(DthCount3_cleaned, 1, function(x) !all(is.na(x))), ]

# Print the modified data frame to verify the changes
head(DthCount3_cleaned)

```

```{r}
#view(DthCount3_cleaned)
```

```{r}
str(DthCount3A_cleaned)
```

```{r, message=FALSE}
#update.packages('mgcv')
library(mgcv)
```

```{r}
# Convert the 'state' column to a factor
DthCount3A_cleaned$State <- as.factor(DthCount3_cleaned$State)

# Display the resulting data frame
head(DthCount3_cleaned)
```

```{r}
# Step 3: Reindex the rows starting from 1
rownames(DthCount3A_cleaned) <- NULL
```

```{r}
str(DthCount3A_cleaned)
```

```{r}
library(ggplot2)

# Bar chart of Death Count by State with skyblue fill and flipped coordinates
ggplot(data = DthCount3A_cleaned, aes(x = State, y = Death_Count)) +
  geom_bar(stat = "identity", fill = "darkblue") +
  labs(title = "Varroa Mites by State",
       x = "State",
       y = "Death Count") +
  theme_minimal() +
  coord_flip() +
  theme(
    axis.text.x = element_text(size = 12),  # Adjust the size as needed
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text.y = element_text(size = 10)
  )

```

# Remove all rows with any NA values

```{r}

library(tidyr)

DeathCnt_data <- DthCount3A_cleaned %>% drop_na()

# Verify that NAs have been removed
glimpse(DeathCnt_data)
```

```{r, message=FALSE, warning=FALSE}
## Example of a bivariate plot between 'state' and 'varroa_mites'
ggplot(DeathCnt_data, aes(x = State, y = Death_Count)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Death Count by State",
       x = "State",
       y = "Death Count")
```

```{r}
DeathCntFinal_data<-DeathCnt_data
```

```{r}
library(ggplot2)

# Bar chart of Death Count by State with skyblue fill and flipped coordinates
ggplot(data = DeathCntFinal_data, aes(x = State, y = Death_Count)) +
  geom_bar(stat = "identity", fill = "darkblue") +
  labs(title = "Varroa Mites by State",
       x = "State",
       y = "Death Count") +
  theme_minimal() +
  coord_flip() +
  theme(
    axis.text.x = element_text(size = 12),  # Adjust the size as needed
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text.y = element_text(size = 10)
  )

```

```{r}
# Write the cleaned data to a new CSV file
write.csv(DeathCntFinal_data, "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/cdcdata-exercise/DeathCntFinal_data.csv", row.names = FALSE)
```

################################Eddie Cruz #################################################################

```{r}
# Install and load necessary packages
library(synthpop)
```

```{r}
# Output cleaned data file to a csv file.
DthCount_data <- read.csv("C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/Storage CDCdata-exercise/DeathCntFinal_data.csv", header = FALSE, stringsAsFactors = FALSE)
```

```{r}
# Load your data (assuming it's saved as a CSV file)
data <- read.csv("DeathCntFinal_data.csv")

```

```{r}
# Subset the data to keep only the first 500 observations
data <- data[1:500, ]
```

```{r}
# Handle missing values (for simplicity, we'll remove rows with any NAs)
data <- na.omit(data)

```

```{r}
# Identify and remove constant columns
constant_columns <- sapply(data, function(col) length(unique(col)) == 1)
data <- data[, !constant_columns]
```

```{r}
# Convert Week_Ending_Date to Date type and create a new variable for month
data$Week_Ending_Date <- as.Date(data$Week_Ending_Date)
```

```{r}
# Remove the original Week_Ending_Date column to avoid issues with high levels
data$Week_Ending_Date <- NULL

```

```{r}
# Convert numeric variables with few levels to factors
data$Year <- as.factor(data$Year)
data$Period <- as.factor(data$Period)

# Create a predictor matrix to handle collinear variables
predictor_matrix <- syn(data)$predictor.matrix
# Set collinear variables to 0 for others' synthesis
predictor_matrix["Period", "Year"] <- 0
predictor_matrix["Year", "Period"] <- 0
predictor_matrix["Avg_No_Deaths", "Death_Count"] <- 0
predictor_matrix["Death_Count", "Avg_No_Deaths"] <- 0

# Generate synthetic data
synth_data <- syn(data, predictor.matrix = predictor_matrix, maxfaclevels = 200, minnumlevels = 5)

# Access the synthetic dataset
synthetic_data <- synth_data$syn

# Save the synthetic dataset to a CSV file
write.csv(synthetic_data, "synthetic_data.csv", row.names = FALSE)
```