### 2024-07-18 Honey breakdown.qmd
This is the update for part 4 project Group10 manuscript date is not 07.17.24

---
title: "Honey Bee Code"
author: "Edward Cruz, Jr."
editor: visual
---

```{r}
# Install necessary packages

# Load the libraries
library(ggplot2)
library(dplyr)
library(scales)
library(showtext)
library(ggimage)
```

```{r}
# Create a data frame with the composition of honey
honey_data <- data.frame(
  component = c("Fructose", "Glucose", "Water", "Maltose", "Trisaccharides, Carbohydrates", "Sucrose", "Minerals, Vitamins, Enzymes"),
  percentage = c(38.5, 31.0, 17.1, 7.2, 4.2, 1.5, 0.5)
)

# Define colors for the segments
honey_data$color <- c("#FFA726", "#FB8C00", "#FFD54F", "#FFB74D", "#90CAF9", "#F06292", "#BA68C8")

# Path to the bee icon
bee_jpg <- "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/DA-6833-02T-202430 Practicum II/bee_2.jpg" # Update with your bee icon path

```

```{r}
ggplot(honey_data, aes(x = "", y = percentage, fill = component)) +
  geom_bar(width = 0.8, stat = "identity", color = "white") +  # Adjust width here
  coord_polar("y", start = 0) +
  scale_fill_manual(values = honey_data$color) +
  geom_text(aes(label = paste0(percentage, "%")), position = position_stack(vjust = 0.6), size = 3, color = "black") +
  labs(
    title = "COMPOSITION OF HONEY",
    subtitle = "Illustration of honey components by percentage",
    fill = NULL
  ) +
  theme_void() +
  theme(
    plot.title = element_text(size = 22, face = "bold", hjust = 0.5, family = "lobster"),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.position = "bottom",
    legend.text = element_text(size = 12)
  ) +
  # Add the bee icon
  ggimage::geom_image(aes(x = 1.5, y = 0, image = bee_jpg), size = 0.1)  # Adjust x and y to position the bee outside the pie chart
```