### 2024-08-02 starter-analysis-exercise/products/report/manuscript_final-2.qmd
removed all r -from view

---
title: "Honey Bee Colonies Impacted by Varroa, American Foulbrood and Global Warming"
subtitle: ""
author: Edward Cruz, Jr., Leonel G. Salazar, Amelia Tristan, Mojisola Popoola
date: today
format:
  html:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../citations.bib 
csl: ../apa.csl
editor: 
  markdown: 
    wrap: 72
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# load a few R packages
library(here)
library(knitr)
library(ggplot2)
library(dplyr)
library(scales)
library(showtext)
library(ggimage)

```

# Summary/Abstract

Group Ten is conducting a comprehensive analysis of historical data from
various public agencies to evaluate the impact of Varroa mites, American
Foulbrood, and global warming on hive losses in the United States.This
research utilizes extensive datasets from the National Agricultural
Statistics Service, the Agricultural Statistics Board, and the United
States Department of Agriculture (USDA), encompassing multiple years of
data. The data highlights the topic on hive losses attributed to mites,
bacterial infections, and environmental factors related to global
warming. Through advanced data visualization techniques in R, we aim to
demonstrate and validate the detrimental effects of these factors on
honeybee colonies, highlighting the consequent implications for honey
production and broader food security. ***Dennis and Kemp's study on
honey bee hive collapse [@hivecollapse] provides important insights into
Allee effects and ecological resilience.*** ***Warmer climate may impact
bees[@2024warmer]***

{{< pagebreak >}}

# Introduction

Honeybees have served as nature's pollinators for centuries. With their
relationship with humans documented as far back as ancient Egyptian and
Hindu cultures. Historically, humans have maintained beehives and
utilized honey for its medicinal properties in various civilizations,
including the Egyptians, Assyrians, Chinese, Greeks, and Romans. Honey's
natural antibacterial qualities made it a valuable treatment for wounds.
This was a practice continued by Romans and Russians during World War I.
Honeybees and other pollinators are essential for food production and
nutritional security. But even though they face numerous survival
challenges Varroa mites pose a significant threat to bee colonies. These
tiny red-brown parasites live on adult honeybees and reproduce on larvae
and pupae. Another major threat is American Foulbrood Disease (AFB).
This is caused by the bacterium Paenibacillus larvae, which is fatal to
honeybee larvae and requires incineration of infected hives to prevent
its spread. The impact of climate change on honeybee colony losses is a
growing area of research. While correlations have been found between
higher winter temperatures and increased colony losses, the effects of
warmer autumn and winter temperatures on colony dynamics and survival
require further investigation. ***Dennis and Kemp's study on honey bee
hive collapse [@hivecollapse] provides important insights into Allee
effects and ecological resilience.*** [@usda_honey_bees;
@usda_nass]\*\*\*

***Warmer climate may impact bees[@2024warmer]***

### Composition of Honey

The first image is a pie chart that illustrates the composition of honey
by percentage. The main components are:

-   **Fructose**: 38.5%
-   **Glucose**: 31%
-   **Water**: 17.1%
-   **Maltose**: 7.2%
-   **Other components**:
    -   Sucrose: 1.5%
    -   Minerals, Vitamins, Enzymes: 0.5%
    -   Trisaccharides, Carbohydrates: 4.2%

This composition indicates that honey is primarily made up of sugars,
specifically fructose and glucose, with water being the third major
component. The presence of minerals, vitamins, and enzymes, though in
smaller amounts, adds nutritional value to honey. This detailed
breakdown underscores honey's role as a natural sweetener with
additional health benefits beyond its primary carbohydrate content.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Create a data frame with the composition of honey
honey_data <- data.frame(
  component = c("Fructose", "Glucose", "Water", "Maltose", "Trisaccharides, Carbohydrates", "Sucrose", "Minerals, Vitamins, Enzymes"),
  percentage = c(38.5, 31.0, 17.1, 7.2, 4.2, 1.5, 0.5)
)

# Define colors for the segments
honey_data$color <- c("#FFA726", "#FB8C00", "#FFD54F", "#FFB74D", "#90CAF9", "#F06292", "#BA68C8")

# Plot with ggplot2
ggplot(honey_data, aes(x = "", y = percentage, fill = component)) +
  geom_bar(width = 0.8, stat = "identity", color = "white") +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = honey_data$color) +
  geom_text(aes(x=1.8, label = paste0(percentage, "%")), position = position_stack(vjust = 0.6), size = 3, color = "black") +
  labs(
    title = "COMPOSITION OF HONEY",
    subtitle = "Illustration of honey components by percentage",
    fill = NULL
  ) +
  theme_void() +
  theme(
    plot.title = element_text(size = 22, face = "bold", hjust = 0.5, family = "lobster"),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.position = "bottom",
    legend.text = element_text(size = 12)
  )
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(sf)
library(maps)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
states <- st_as_sf(map("state", plot = FALSE, fill = TRUE))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
varroa_data <- data.frame(
  state = c("alabama", "alaska", "arizona", "arkansas", "california", "colorado", 
            "connecticut", "delaware", "florida", "georgia", "hawaii", "idaho", 
            "illinois", "indiana", "iowa", "kansas", "kentucky", "louisiana", 
            "maine", "maryland", "massachusetts", "michigan", "minnesota", 
            "mississippi", "missouri", "montana", "nebraska", "nevada", "new hampshire", 
            "new jersey", "new mexico", "new york", "north carolina", "north dakota", 
            "ohio", "oklahoma", "oregon", "pennsylvania", "rhode island", 
            "south carolina", "south dakota", "tennessee", "texas", "utah", 
            "vermont", "virginia", "washington", "west virginia", "wisconsin", "wyoming"),
  year = c("1986 - 1987", "1988 - 1989", "1990 - 1991", "1992 - 1993", "1994 - 1995",
           "1986 - 1987", "1988 - 1989", "1990 - 1991", "1992 - 1993", "1994 - 1995",
           "1986 - 1987", "1988 - 1989", "1990 - 1991", "1992 - 1993", "1994 - 1995",
           "1986 - 1987", "1988 - 1989", "1990 - 1991", "1992 - 1993", "1994 - 1995",
           "1986 - 1987", "1988 - 1989", "1990 - 1991", "1992 - 1993", "1994 - 1995",
           "1986 - 1987", "1988 - 1989", "1990 - 1991", "1992 - 1993", "1994 - 1995",
           "1986 - 1987", "1988 - 1989", "1990 - 1991", "1992 - 1993", "1994 - 1995",
           "1986 - 1987", "1988 - 1989", "1990 - 1991", "1992 - 1993", "1994 - 1995",
           "1986 - 1987", "1988 - 1989", "1990 - 1991", "1992 - 1993", "1994 - 1995",
           "1986 - 1987", "1988 - 1989", "1990 - 1991", "1992 - 1993", "1994 - 1995")
)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
states <- merge(states, varroa_data, by.x = "ID", by.y = "state", all.x = TRUE)
```

### Spread of Varroa Mites

The second image is a map illustrating the spread of Varroa mites in the
United States from 1986 to 1995. The map uses different colors to
represent the year of the first appearance of Varroa mites in each
state:

-   **1986 - 1987**: Red
-   **1988 - 1989**: Yellow
-   **1990 - 1991**: Green
-   **1992 - 1993**: Blue
-   **1994 - 1995**: Grey
-   **NA**: States where data is not available

The map reveals the progressive spread of Varroa mites from the
mid-1980s to the mid-1990s, affecting bee colonies across the country.
The earliest appearances were concentrated in specific regions and
gradually spread to more states over time. This visual representation
highlights the increasing geographical distribution of Varroa mites and
underscores the growing challenge they pose to beekeeping and
agricultural industries in the United States.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = states) +
  geom_sf(aes(fill = year), color = "#292929", size = 0.2) +
  scale_fill_manual(
    values = c(
      "1986 - 1987" = "#CD2626", 
      "1988 - 1989" = "#CD9B1D", 
      "1990 - 1991" = "#FFD700", 
      "1992 - 1993" = "#00CD00", 
      "1994 - 1995" = "#104E8B"
    ),
    na.value = "#757575"
  ) +
  theme_minimal() +
  labs(
    title = "Spread of Varroa Mites",
    fill = "Year of First Appearance"
  )

```

[@varroamap] ***Varroa Map of infestation periods***

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load necessary libraries
library(tidyverse)

# Create a data frame with the provided data
data <- tibble(
  Year = 1901:2023,
  `Earth's surface` = c(-0.15, -0.43, -1.4, -0.86, -1.02, -0.29, -0.54, 0.06, -0.59, 0.4, 0.01, -1.79, -0.48, -0.18, -0.57, -1.17, -1.96, -0.15, -0.47, -0.95, 1.78, 0.01, -0.38, -1.43, 0.5, -0.07, 0.13, -0.1, -1.17, -0.04, 1.52, -0.29, 0.97, 2.08, -0.12, 0.13, -0.47, 1.16, 1.24, -0.13, 0.64, -0.18, 0.05, -0.19, -0.27, 0.93, -0.1, -0.41, 0, -0.63, -0.9, 0.25, 1.35, 1.31, -0.33, 0.32, 0.02, -0.09, 0.09, -0.58, -0.15, -0.12, 0.24, -0.35, -0.33, -0.53, -0.26, -0.7, -0.52, -0.41, -0.36, -0.65, 0.27, 0.24, -0.52, -0.55, 0.53, -0.97, -1.14, 0.37, 1.1, -0.67, -0.14, -0.04, -0.72, 1.3, 1.31, 0.61, -0.18, 1.49, 1.14, 0.58, -0.76, 0.85, 0.63, -0.13, 0.18, 2.21, 1.86, 1.25, 1.68, 1.19, 1.24, 1.08, 1.62, 2.23, 1.63, 0.27, 0.37, 0.96, 1.16, 3.26, 0.41, 0.52, 2.38, 2.9, 2.53, 1.5, 0.66, 2.35, 2.49, 1.37, 2.39),
  `Lower troposphere (UAH)` = c(rep(NA, 78), 0.049093434, 0.731593434, 1.102093434, 0.008593434, 0.034093434, 0.292093434, 0.110593434, 1.220593434, 1.054093434, 0.698593434, 0.548593434, 1.451593434, 1.031593434, 0.484093434, -0.174406566, 1.043593434, 0.797593434, 0.502093434, 0.457093434, 1.691593434, 1.796593434, 1.351093434, 1.255093434, 0.994093434, 1.223593434, 0.769093434, 1.303093434, 1.444093434, 1.765093434, 0.467593434, 0.524593434, 0.929593434, 1.013593434, 1.955593434, 0.655093434, 0.763093434, 2.234593434, 2.030593434, 2.252593434, 1.645093434, 1.133593434, 1.970593434, 2.059093434, 1.567093434, NA),
  `Lower troposphere (RSS)` = c(rep(NA, 78), -0.808971111, 0.178178889, 0.658778889, -0.579771111, -0.383871111, -0.217371111, -1.053021111, 0.737978889, 1.111328889, 0.557978889, 0.221828889, 1.280228889, 1.043228889, 0.585278889, -0.261321111, 1.078178889, 0.772628889, 0.278078889, 0.418178889, 1.911578889, 1.882928889, 1.471328889, 1.536578889, 1.144628889, 1.465628889, 1.058528889, 1.583078889, 1.749278889, 1.858928889, 0.503378889, 0.595928889, 0.997028889, 1.109378889, 2.677328889, 0.696278889, 0.781478889, 2.547878889, 2.476328889, 2.350478889, 1.793978889, 1.221428889, 2.290628889, 2.485328889, 1.596878889, 1.886228889)
)
# Pivot the data for plotting
data_long <- data %>%
  pivot_longer(cols = c(`Earth's surface`, `Lower troposphere (UAH)`, `Lower troposphere (RSS)`), 
               names_to = "Measurement", values_to = "Temperature")

# Create the plot
ggplot(data_long, aes(x = Year, y = Temperature, color = Measurement)) +
  geom_line(size = 1) +
  geom_point(size = 1.5) +
  theme_minimal() +
  labs(
    title = "Temperature Changes Over the Years",
    x = "Year",
    y = "Temperature Anomaly (°C)",
    color = "Measurement"
  ) +
  scale_color_manual(values = c("Earth's surface" = "#254976", "Lower troposphere (UAH)" = "#4A8C1D", "Lower troposphere (RSS)" = "#B5153A"))

```

[@epa_climate] ***Global warming trends from 1900-2024***

***Spread of Varroa Mite by Year of First Apperance [@varroamap]***
Written By:sdns6mchl4. (2016, February 24). Varroa mite spread in the
United States. Beesource Beekeeping Forums.
https://www.beesource.com/threads/varroa-mite-spread-in-the-united-states.365462/

\## General Background Information

### **Uncapped Honey Floresville,Texas Hive**

![](/images/Bee%20Frame%202024.jpg){fig-align="center"}

### **Capped Honey one Month Later same hive frame- Italian bees Floresville, Texas**

![](/images/Bee%20Frame%20Capped%20Honey.jpg){fig-align="center"}

## Description of data and data source

Bee colonies maintained by beekeepers are considered livestock by the
USDA due to their ability to produce honey, a consumable food item, and
their essential role in assisting farmers with pollination crop seasons.
Given the importance of bee colonies in agriculture, it was logical to
source data from the following two authoritative websites: 1. USDA
National Agricultural Statistics Service (NASS): This site provides
comprehensive agricultural data, including statistics on honey
production and colony health. 2. Bee Informed Partnership: This site
offers detailed insights and research on bee colony management and
health, contributing valuable information on the status and trends of
bee populations. [@usda_honey_bees]

Index Catalog // USDA Economics, Statistics and Market Information
System. (n.d.-a).
https://usda.library.cornell.edu/catalog?f%5Bkeywords_sim%5D%5B%5D=honey+bees&locale=en\
USDA - National Agricultural Statistics Service - Surveys - honey bee
surveys and reports. (n.d.).
https://www.nass.usda.gov/Surveys/Guide_to_NASS_Surveys/Bee_and_Honey/

## Questions/Hypotheses to be addressed

Hypothesis: "The negative impacts of mites, bacterium, and global
warming have detrimental effects on honeybee colonies in the United
States and Texas, which in turn will lead to a decline in honey
production and negatively impact food production." This hypothesis can
be tested and validated through a visualization of outcomes using R,
demonstrating the relationship between these factors and their effects
on honeybee colonies.

### **Bacterium Infection Foul Brood**

![](/images/American-Foul-Brood-Comb-1024x768-1.jpg){fig-align="center"}

### **Dead bees resulting from extreme heat found in hive**

![](/images/dead-bees-hive-colony-collapse-600nw-2289939105.webp)
{{< pagebreak >}}

# Methods

*Describe your methods. That should describe the data, the cleaning
processes, and the analysis approaches. You might want to provide a
shorter description here and all the details in the supplement.*

## Schematic of workflow

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Install necessary packages
if (!require("DiagrammeR")) install.packages("DiagrammeR")

# Load the DiagrammeR package
library(DiagrammeR)

# Create the flowchart using DiagrammeR
grViz("
  digraph Data_Analysis_Flowchart {
    graph [layout = dot, rankdir = LR, label = 'Workflow Schematic', fontsize = 45, labelloc = 't']

    node [style = filled, fontsize = 30, width = 2, height = 2]

    Define_Objectives [label = 'Define Objectives', shape = oval, color = lightblue]
    Collect_Data [label = 'Collect Data', shape = box, color = lightgrey]
    Clean_Data [label = 'Clean Data', shape = ellipse, color = lightyellow]
    Analyze_Data [label = 'Analyze Data', shape = diamond, color = lightcoral]
    Interpret_Results [label = 'Interpret Results', shape = trapezium, color = lightgrey]
    Visualize_Data [label = 'Visualize Data', shape = parallelogram, color = lightpink]
    Compile_Report [label = 'Compile Report', shape = house, color = lightcyan]
    Review_Revise [label = 'Review & Revise', shape = invhouse, color = lightgoldenrodyellow]
    Present_Report [label = 'Present Report', shape = hexagon, color = lightseagreen]

    Define_Objectives -> Collect_Data
    Collect_Data -> Clean_Data
    Clean_Data -> Analyze_Data
    Analyze_Data -> Interpret_Results
    Interpret_Results -> Visualize_Data
    Visualize_Data -> Compile_Report
    Compile_Report -> Review_Revise
    Review_Revise -> Present_Report
  }
")

```

## Data aquisition

Data was derived from the United States Department of Agriculture (USDA) website, which permits data sets to be downloaded in CSV(comma-separated values) format.

## Data import and cleaning

We decided to clean out our data from a few different datasets. We had
to remove blank spaces and columns that were not pertinent to our
analysis. We then filtered out other observations that did not directly
deal with the data we are exploring. We are looking for cause of death
to bee colonies and how they are affected by mites and climate change so
we wanted to single out data that represented the losses so we can
explore the different states by year and determine how the colonies were
affected.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
library(ggplot2)
library(knitr)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}

library(here)

# Install and load the here package
#install.packages("here")
library(here)

here()
# Create a 'data' directory within the project
dir.create(here("data"))

# Read the CSV file
data <- read.csv(here("Databystate.csv"))

#data <- read.csv("C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II #Repository/P2-Practicum-II-Portfolio-EdwardCruz/DatabyState.csv")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Select all columns except 3, 6, and 9
Data_Clean <- dplyr::select(data, -c(3, 6, 9))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Output cleaned data file to a csv file.

write.csv(Data_Clean, "Data_Clean.csv", row.names = FALSE)

#write.csv(Data_Clean, "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II #Repository/P2-Practicum-II-Portfolio-EdwardCruz/docs/Databystate_Clean.csv")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Assuming your data frame is named "data"

# Filter for rows where "Loss" or "Deadout" is present in any column (case-insensitive)

library(stringr)  # Load stringr package for regular expressions
data_filtered <- Data_Clean[rowSums(sapply(data, grepl, pattern = c("Loss"), ignore.case = TRUE)) > 0, ]
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Output cleaned data file to a csv file.

write.csv(data_filtered, "Data_Filtered.csv", row.names = FALSE)

#write.csv(Data_Clean, "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II #Repository/P2-Practicum-II-Portfolio-EdwardCruz/docs/Databystate_Filtered.csv")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# View data str to see what class they are.

str(data_filtered)
```

### Boxplot of Value by Year and Data Item

The third image is a boxplot comparing the values for "LOSS, COLONY
COLLAPSE DISORDER" and "LOSS, DEADOUT" from 2015 to 2017. The boxplot
provides insights into: - The central tendency and spread of losses due
to both disorders. - Presence of outliers indicates that there are
occasional, significantly higher losses. - Comparative analysis shows
that both types of losses have varying distributions each year.

This indicates that while both types of losses are present each year,
the magnitude and frequency of extreme loss events can differ,
suggesting variability in the factors contributing to colony losses.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Load necessary libraries
library(ggplot2)
library(here)

str(data_filtered)

# Load the dataset
data_filtered <- read.csv(here::here("Data_Filtered.csv"), stringsAsFactors = FALSE)

# Convert 'Value' to numeric
data_filtered$Value <- as.numeric(gsub(",", "", data_filtered$Value))

# Handle missing values by replacing NAs with the median value
data_filtered$Value[is.na(data_filtered$Value)] <- median(data_filtered$Value, na.rm = TRUE)

# Boxplot of Value by Year and Data.Item
ggplot(data_filtered, aes(x = factor(Year), y = Value, fill = Data.Item)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Boxplot of Value by Year and Data.Item", x = "Year", y = "Value") +
  theme_minimal()
```

### Scatterplot of Value by State ANSI

The fourth image is a scatterplot showing values for different states
(ANSI coded) across the years 2015 to 2017. Key points include: - Each
color represents a different year (2015: Red, 2016: Green, 2017:
Blue). - The scatterplot reveals the distribution of values across
states and highlights any trends or patterns over the three years.

This scatterplot indicates that while some states experience
consistently high or low values, others show significant variability
year over year. This could suggest regional differences in factors
affecting bee colony health.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Scatterplot of Value by State.ANSI
ggplot(data_filtered, aes(x = State.ANSI, y = Value, color = factor(Year))) +
  geom_point() +
  labs(title = "Scatterplot of Value by State.ANSI", x = "State.ANSI", y = "Value") +
  theme_minimal()

```

### Boxplot of Value by Period and Data Item

The fifth image is a boxplot comparing the values for "LOSS, COLONY
COLLAPSE DISORDER" and "LOSS, DEADOUT" across different periods of the
year: - **APR THRU JUN** - **JAN THRU MAR** - **JUL THRU SEP** - **OCT
THRU DEC**

Key insights include: - Seasonal trends in colony losses, with some
periods showing higher median losses or more variability. - Outliers
indicate extreme loss events occurring in specific periods.

This seasonal analysis suggests that certain times of the year are more
prone to higher losses, potentially due to environmental factors,
resource availability, or management practices.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load necessary library
library(ggplot2)
library(scales)

# Assuming data_filtered is already loaded and has the required columns

# Create the plot
p <- ggplot(data_filtered, aes(x = Period, y = Value, fill = Data.Item)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Boxplot of Value by Period and Data.Item", x = "Period", y = "Value")

# Print the plot
print(p)

```

### Boxplot of Value by Period and Data Item

The fifth image is a boxplot comparing the values for "LOSS, COLONY
COLLAPSE DISORDER" and "LOSS, DEADOUT" across different periods of the
year: - **APR THRU JUN** - **JAN THRU MAR** - **JUL THRU SEP** - **OCT
THRU DEC**

Key insights include: - Seasonal trends in colony losses, with some
periods showing higher median losses or more variability. - Outliers
indicate extreme loss events occurring in specific periods.

This seasonal analysis suggests that certain times of the year are more
prone to higher losses, potentially due to environmental factors,
resource availability, or management practices.

### Overall Analysis

-   The composition of honey indicates its high sugar content, essential
    for energy but also highlighting the presence of beneficial
    components like enzymes and vitamins.
-   The spread of Varroa mites map highlights the widespread and growing
    impact of these pests over time, emphasizing the need for ongoing
    management and control efforts.
-   The boxplots and scatterplots provide a detailed view of the
    variability and trends in colony losses due to different factors,
    both over time and across different regions. These plots suggest
    that while losses are a consistent issue, their magnitude and causes
    can vary widely, pointing to the need for tailored strategies to
    address colony health.

These visualizations collectively provide a comprehensive overview of
the challenges faced by bee colonies, from composition and health to
external threats like Varroa mites and seasonal variations in losses.

### **VARROA MITE EXPOSED**

![](/images/_20220406_on_varroadestructor_minden_00463080.jpg){fig-align="center"}

### **VARROA MITE**

![](/images/Varrpa%201.webp){fig-align="center" width="527"}

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(reshape2)
library(cluster)
library(car)
library(factoextra)
library(rpart)
library(rpart.plot)
library(randomForest)
library(e1071)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load your dataset
# Assuming your data frame is named 'data_hcny'

data_hcny <- read.csv(here("hcny_CleanDraft.csv"))

#data_hcny <- read.csv("C:/Users/Leonel/Desktop/MSDA/MS Data Analytics/Current Class/DA 6833/Practicum 2 Github/leonelsalazar-P2-portfolio/Group Project/hcnydata_cleaned.csv")

data_clean <- na.omit(data_hcny)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Descriptive Statistics
print("Descriptive Statistics:")
summary(data_clean)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
str(data_clean)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Visualizations

# Bar Plot for varroa_mites by State
ggplot(data_clean, aes(x = state, y = varroa_mites)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Varroa Mites by State", y = "Percentage", x = "State")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
str(data_clean)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Stacked Bar Plot
data_long <- data_clean %>%
  gather(key = "factor", value = "percentage", -state)

ggplot(data_long, aes(x = state, y = percentage, fill = factor)) +
  geom_bar(stat = "identity") +
 theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Factors Affecting Bees by State", y = "Percentage", x = "State")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Heatmap
data_melt <- melt(data_clean, id.vars = "state")

ggplot(data_melt, aes(x = variable, y = state, fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "red") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Heatmap of Factors Affecting Bees", y = "State", x = "Factor")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Pie Chart for the United States
us_data <- data_clean %>%
  filter(state == "United States") %>%
  dplyr::select(-state) %>%
  gather(key = "factor", value = "percentage")

ggplot(us_data, aes(x = "", y = percentage, fill = factor)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y") +
  theme_void() +
  labs(title = "Distribution of Factors Affecting Bees in the United States")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Correlation Analysis
cor_data <- cor(data_clean[,-1])
print("Correlation Matrix:")
print(cor_data)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Assumption Checks for Descriptive Analysis
print("Assumption Checks:")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Normality check
shapiro_results <- lapply(data_clean[-1], function(column) shapiro.test(column))
print("Shapiro-Wilk Test for Normality:")
print(shapiro_results)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Homogeneity of variances

levene_results <- leveneTest(varroa_mites ~ state, data = data_clean)
print("Levene's Test for Homogeneity of Variances:")
print(levene_results)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Cluster Analysis
data_cluster <- data_clean %>%
  filter(state != "United States") %>%
  dplyr::select(-state)

hc <- hclust(dist(data_cluster), method = "complete")
plot(hc, labels = data_clean$state[data_clean$state != "United States"], main = "Hierarchical Clustering of States")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Principal Component Analysis (PCA)
data_pca <- data_clean %>%
  filter(state != "United States") %>%
  dplyr::select(-state) %>%
  scale()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
pca_result <- prcomp(data_pca, center = TRUE, scale. = TRUE)
summary(pca_result)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
fviz_pca_var(pca_result, col.var = "contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"))

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
str(data_clean)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Support Vector Machine (SVM)
svm_model <- svm(varroa_mites ~ other_pests + disease + pesticides + other + unknown, data = data_clean)
summary(svm_model)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Assuming predictions and data_clean$varroa_mites are already created

# Example data
data_clean <- data.frame(
  Varroa_Mites = c(5, 10, 15, 20),
  other_variable = c(2, 4, 6, 8)
)

# Fit a linear model
model <- lm(Varroa_Mites ~ other_variable, data = data_clean)

# Generate predictions
predictions <- predict(model, newdata = data_clean)

# Plot the data
plot(predictions, data_clean$varroa_mites, 
     xlab = "Predictions", 
     ylab = "Varroa Mites", 
     main = "Predictions vs Varroa Mites",
     las = 1, # makes y-axis labels horizontal
     cex.lab = 1.2, # increases axis labels size
     cex.axis = 1.2) # increases tick labels size

```

### **Honey Bee Collecting Pollen**

![](/images/Bee%20Feeding.jpg){fig-align="center" width="313"}

### Honey Bee with exposed bloated Varroa Mite

![](images/varroa-mite-bee.jpg){fig-align="center"}

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Write the cleaned data to a new CSV file

# Assuming your dataset is named data_hcny
data_hcny_clean <- data_hcny %>% drop_na()

write.csv(data_hcny_clean, here("hcnydata_cleaned.csv"), row.names = FALSE)

#write.csv(hcnydata_cleaned, "C:/Users/ecruz/OneDrive/Documents/UTSA - Data Science Program/Semester Classes/Practicum II Repository/P2-Practicum-II-Portfolio-EdwardCruz/hcny_CleanDraft.csv", row.names = FALSE)
```

## Full analysis

##The full analysis is still pending###

{{< pagebreak >}}

# References